---
title: "Homework #1: Baseball Analysis"
subtitle: "Data 621 Business Analytics and Data Mining"
author: "Aadi Kalloo, Nathan Lim, Asher Meyers, Daniel Smilowitz, Logan Thomson"
date: "Due June 19, 2016"
output: pdf_document
geometry: margin=0.5in
---

```{r, echo = FALSE}
library(stringr)
library(pander)
library(knitr)
library(ggplot2)
library(gridExtra)

trainingdata = read.csv("https://raw.githubusercontent.com/aadikalloo/AadiMSDA/master/IS621-Data-Mining/moneyball-training-data.csv")
####### Mean and Medians Table
mean_median_df = data.frame(matrix(0, nrow = 17, ncol = 2))

mean2 <- function(x)   {mean(x, na.rm = TRUE)}
median2 <- function(x) {median(x, na.rm = TRUE)}

means = as.data.frame(lapply(trainingdata, mean2))
medians = as.data.frame(lapply(trainingdata, median2))
lengths = as.data.frame(lapply(trainingdata, length))

mean_median_df[, 1] = names(means)
mean_median_df[, 2] = t(means[1, ])
mean_median_df[, 3] = t(medians[1, ])
#mean_median_df[, 4] = t(lengths[1, ])

names(mean_median_df) = c("VAR_NAME", "MEAN", "MEDIAN")
#kable(mean_median_df, digits = 2)
####################################################

########Correlations to Wins

cor_df = data.frame(matrix(0, nrow = 15, ncol = 2))

cors = as.data.frame(cor(trainingdata$TARGET_WINS, trainingdata[, 3:17], use = "pairwise.complete.obs"))
cor_df[, 1] = names(cors)
cor_df[, 2] = t(cors[1, ])

names(cor_df) = c("VAR_NAME", "CORRELATION TO WINS (r)")

#kable(cor_df, digits = 2)
####################################################

########Missing Values per variable
mv_df = data.frame(matrix(0, nrow = 17, ncol = 2))

num_missing <- function(x)   {sum(is.na(x))}


missingvalues = as.data.frame(lapply(trainingdata, num_missing))
mv_df[, 1] = names(missingvalues)
mv_df[, 2] = t(missingvalues[1, ])

names(mv_df) = c("VAR_NAME", "NUM_MISSING")

#kable(mv_df, digits = 2)
####################################################


data_exp = merge(mean_median_df, cor_df, by.x = "VAR_NAME", by.y = "VAR_NAME")
data_exp = merge(data_exp, mv_df, by.x = "VAR_NAME", by.y = "VAR_NAME")
temp = as.data.frame(cbind(mean_median_df[2,], NA, NA))
names(temp) = names(data_exp)
data_exp = rbind(temp, data_exp)
```



#Data Exploration

The data analyzed in this report includes 2276 professional baseball teams for the years 1871-2006. In total, 16 variables were present in the data provided. Included below is a summary of descriptive statistics, correlations to wins, and the number of missing values for each variable in the provided data set:

```{r, echo = FALSE}
kable(data_exp)
```

Below are graphs that show the relationship to _Target Wins_ for the three variables with the highest correlation coefficient: 

```{r, echo = FALSE}
par(mfrow = c(1, 3))
top3correlations = c(0,0,7,4,0,0,6)
for (plot_count in c(3, 4, 7)) {
  plot(x = trainingdata[, plot_count], y = trainingdata$TARGET_WINS, xlab = names(trainingdata)[plot_count], ylab = "Target Wins", main = paste0(names(trainingdata)[plot_count], " r = ", round(data_exp[top3correlations[plot_count], 4], 3) ))
}
```

The full array of correlations graphs may be found in Appendix A. 
<!-- -->

#Data Preparation

It was determined that the _Hits By Pitch_ variable had too many missing values to be useful for regression, and thus this variable was excluded from the model building process.  

#Model Creation

#####Load Data

```{r, echo = FALSE}
trainingDataRaw = trainingdata
data_no_index <- trainingDataRaw[,c(2:17)] # delete index
```


#####Imputing Missing values with median
```{r}
for (i in 1:16){
data_no_index[,i][is.na(data_no_index[,i])] <-  median(data_no_index[,i], na.rm = TRUE)
}
df_new=data_no_index
summary(df_new)
```


####Use all the variables to see p value of each variables.
```{r}
fit_all <- lm(TARGET_WINS ~ . , df_new)
summary(fit_all)
```









####Model --(Nathan)

```{r}

g1 <- ggplot(df_new, aes(x=TEAM_FIELDING_E)) + geom_histogram(binwidth = 10)
g2 <- ggplot(df_new, aes(x=log(TEAM_FIELDING_E))) + geom_histogram(binwidth = 0.05)
grid.arrange(g1, g2, ncol=2)

model2<- lm(TARGET_WINS ~  TEAM_BATTING_H + TEAM_BASERUN_SB + TEAM_FIELDING_DP + 
              log(TEAM_FIELDING_E), df_new)

par(mfrow=c(2,2)); plot(model2)
par(mfrow=c(1,1)); plot(model2$residuals)
summary(model2)

plot(fitted(model2), residuals(model2))
abline(h=0, col='red')

```
#####Model Summary Table

|Model #|# of Predictors|Adj. R^2|F-Statistic|P-Value|Residual Standard Error|Degrees of Freedom|
|---|---|---|---|---|---|---|
|1|11|0.31|95|2.2e-16|13.07|2264


#####Model 1: Simple Full Linear Regression, With Removal of Non-Significant Predictors



```{r}
dfraw <- read.csv(url("https://raw.githubusercontent.com/dsmilo/DATA621/master/HW1/data/moneyball-training-data.csv"))
data_no_index <- dfraw[ ,c(2:17)] #Remove INDEX column
for (i in 1:16) {
  data_no_index[,i][is.na(data_no_index[,i])] <-  median(data_no_index[,i], na.rm = TRUE)
}
df_new <- data_no_index
summary(df_new)


fit <- lm(TARGET_WINS~., df_new)
summary(fit)
fit <- update(fit, .~.-TEAM_PITCHING_BB)
summary(fit)
fit <- update(fit, .~.-TEAM_BATTING_HBP)
summary(fit)
fit <- update(fit, .~.-TEAM_BASERUN_CS)
summary(fit)
fit <- update(fit, .~.-TEAM_PITCHING_HR)
summary(fit)
```

#####Model 2: SLR Bounded by Recent MLB Data (1962-)

```{r}
dfraw <- read.csv(url("https://raw.githubusercontent.com/dsmilo/DATA621/master/HW1/data/moneyball-training-data.csv"))
dfremove <- subset(dfraw, INDEX == 1347)$INDEX 
df <- subset(dfraw, !(INDEX %in% dfremove))
data_no_index <- df[ ,c(2:17)] #Remove INDEX column
for (i in 1:16) {
  data_no_index[,i][is.na(data_no_index[,i])] <-  median(data_no_index[,i], na.rm = TRUE)
}
df <- data_no_index
MLB <- read.csv(url("https://raw.githubusercontent.com/dsmilo/DATA621/master/HW1/data/MLB.Stats.1962-2015.csv"))

df$TEAM_BATTING_BB <- ifelse(df$TEAM_BATTING_BB < min(MLB$BB) | df$TEAM_BATTING_BB > max(MLB$BB), median(df$TEAM_BATTING_BB), df$TEAM_BATTING_BB)
df$TEAM_BATTING_H <- ifelse(df$TEAM_BATTING_H < min(MLB$H) | df$TEAM_BATTING_H > max(MLB$H), median(df$TEAM_BATTING_H), df$TEAM_BATTING_H)
df$TEAM_BATTING_2B <- ifelse(df$TEAM_BATTING_2B < min(MLB$X2B) | df$TEAM_BATTING_2B > max(MLB$X2B), median(df$TEAM_BATTING_2B), df$TEAM_BATTING_2B)
df$TEAM_BATTING_3B <- ifelse(df$TEAM_BATTING_3B < min(MLB$X3B) | df$TEAM_BATTING_3B > max(MLB$X3B), median(df$TEAM_BATTING_3B), df$TEAM_BATTING_3B)
df$TEAM_BATTING_HR <- ifelse(df$TEAM_BATTING_HR < min(MLB$HR) | df$TEAM_BATTING_HR > max(MLB$HR), median(df$TEAM_BATTING_HR), df$TEAM_BATTING_HR)
df$TEAM_BATTING_SO <- ifelse(df$TEAM_BATTING_SO < min(MLB$SO) | df$TEAM_BATTING_SO > max(MLB$SO), median(df$TEAM_BATTING_SO), df$TEAM_BATTING_SO)
df$TEAM_PITCHING_SO <- ifelse(df$TEAM_PITCHING_SO < min(MLB$SO.1) | df$TEAM_PITCHING_SO > max(MLB$SO.1), median(df$TEAM_PITCHING_SO), df$TEAM_PITCHING_SO)
df$TEAM_BASERUN_SB <- ifelse(df$TEAM_BASERUN_SB < min(MLB$SB) | df$TEAM_BASERUN_SB > max(MLB$SB), median(df$TEAM_BASERUN_SB), df$TEAM_BASERUN_SB)

fit <- lm(TARGET_WINS~.,df)
summary(fit)
fit <- update(fit, .~.-TEAM_PITCHING_SO)
summary(fit)
fit <- update(fit, .~.-TEAM_BATTING_3B)
summary(fit)
fit <- update(fit, .~.-TEAM_BATTING_HBP)
summary(fit)
fit <- update(fit, .~.-TEAM_PITCHING_BB)
summary(fit)
fit <- update(fit, .~.-TEAM_BASERUN_CS)
summary(fit)
fit <- update(fit, .~.-TEAM_PITCHING_H)
summary(fit)
fit <- update(fit, .~.-TEAM_BATTING_HR)
summary(fit)
```

```{r}
#dfraw <- read.csv(url("https://raw.githubusercontent.com/dsmilo/DATA621/master/HW1/data/moneyball-training-data.csv"))
dfraw = trainingdata
dfremove <- subset(dfraw, TEAM_BATTING_SO == 0 | TEAM_PITCHING_SO == 0 | TEAM_BASERUN_SB == 0 | TEAM_BATTING_BB == 0 | TEAM_PITCHING_SO > 2500 | TEAM_PITCHING_BB == 0 )$INDEX #Identify rows for removal, have zero strikeouts or bases stolen
df <- subset(dfraw, !(INDEX %in% dfremove))
head(df)
df1 <- df[, -c(1,10,11)] #Remove caught stealing and hit by pitcher variables
#View(df)
#View(df1)
#summary(df)

df$TEAM_BATTING_HSO <- df$TEAM_BATTING_H/df$TEAM_BATTING_SO #Ratio of hits to strikeouts

fit1 <- lm(TARGET_WINS~.-TEAM_PITCHING_HR-TEAM_BATTING_SO-TEAM_BATTING_H, df)#Non-significant predictors removed
summary(fit1)
step1 <- step(fit1)
summary(step1)

#Correlation Matrix
#View(round(cor(df1),2))

#These are variables that I tried but didn't turn out to be valuable
df1$TEAM_BATTING_1B <- df1$TEAM_BATTING_H - df1$TEAM_BATTING_2B - df1$TEAM_BATTING_3B - df1$TEAM_BATTING_HR #Singles - 1st Base Hits
df1$TEAM_BATTING_HRP <- df1$TEAM_BATTING_HR/df1$TEAM_BATTING_H #Home runs as a percentage of base hits
```


#####Create a linear model using all predictors. The INDEX column is excluded.
```{r}
FullModel <- lm(TARGET_WINS ~.-INDEX, trainingdata)
summary(FullModel) #Summary of full model
```


#####Put full model through stepwise regression, where predictors with less significance are sequentially removed.
```{r}
stepFull <- step(FullModel) 
summary(stepFull)

```


```{r}
#####Generate predictions using the stepFull model
predictionsStepFull <- predict(stepFull, trainingDataRaw) 
#View(predictionsStepFull)
```

#####Generate the RMSE of the stepFull model
```{r}
rmseStep <- sqrt(mean((trainingDataRaw$TARGET_WINS[!is.na(predictionsStepFull)] - predictionsStepFull[!is.na(predictionsStepFull)])^2)) 
rmseStep
```

```{r}
par(mfrow=c(2,2)) #Set up a four panel plot for evaluating regression
plot(stepFull) #Displays Residuals vs Fitted, Scale-Location,  and Normal Q-Q.
```

#####Evaluation of Stepwise model without TEAM_BATTING_HBP
```{r}
trainingDataRaw = trainingdata
ReducedModel <- lm(TARGET_WINS ~., trainingDataRaw[,c(2:10, 12:17)])
summary(ReducedModel)
stepReduced <- step(ReducedModel)
predictionsStepReduced <- predict(stepReduced, trainingDataRaw[,c(2:10, 12:17)])
rmseStepR <- sqrt(mean((trainingDataRaw$TARGET_WINS[!is.na(predictionsStepReduced)] - predictionsStepReduced[!is.na(predictionsStepReduced)])^2))
rmseStepR
```

#Model Selection and Prediction
