---
title: "HW2_nathanlim"
author: "NathanLim"
date: "June 19, 2016"
output: html_document
---

2. Use the table() function to get the raw confusion matrix for this scored dataset.
```{r}
library(dplyr)
df <- read.csv(url("https://raw.githubusercontent.com/dsmilo/DATA621/master/HW2/Data/classification-output-data.csv"))
head(df)
(confusion <- table(select(df, class, scored.class)))
predicted="scored.class"
```

3. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified,
and returns the accuracy of the predictions.

```{r}
accuracy_function <- function(df, scored.class){
  confusion <- table(select(df, class, scored.class))
  accuracy <- (confusion[1,1] + confusion[2,2])/sum(confusion)
  return(accuracy)
}

accuracy_function(df,scored.class)
```

4. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified,
and returns the classification error rate of the predictions.


```{r}
CER_function<- function(df, predicted){
  confusion <- table(select(df, class, get(predicted)))
  classification_error_rate <- (confusion[1,2] + confusion[2,1])/sum(confusion)
  return(classification_error_rate)
}

CER_function(df,predicted)
```

Verify by summing the rates
```{r}
accuracy_function(df,predicted) +  CER_function(df,predicted)
```


5. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the precision of the predictions.


```{r}
precision_function <- function(df, predicted){
  confusion <- table(select(df, class, get(predicted)))
  precision <- confusion[2,2]/(confusion[1,2] + confusion[2,2])
  return(precision)
}

precision_function(df,predicted)
```


6. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified,
and returns the sensitivity of the predictions. Sensitivity is also known as recall.



```{r}

sensitivity_function <- function(df, predicted){
  confusion <- table(select(df, class, get(predicted)))
  if(dim(confusion)[2] == 1){
      if(colnames(confusion) == 0){
          sensitivity <- 0
           } else{
          sensitivity <- 1
             }
  } else{
      sensitivity <- confusion[2,2]/(confusion[2,1] + confusion[2,2])
  }

  return(sensitivity)
}

sensitivity_function (df, predicted)
```


7. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified,
and returns the specificity of the predictions.

```{r}
specificity_function <- function(df,predicted){
  confusion <- table(select(df, class, get(predicted)))  
  if(dim(confusion)[2] == 1){
    if(colnames(confusion) == 0){
      specificity <- 1
    } else{
      specificity <- 0
    }
  } else{
    specificity <- confusion[1,1]/(confusion[1,1] + confusion[1,2])
  }
  return(specificity)
}

specificity_function(df, predicted)
```

8. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified,
and returns the F1 score of the predictions.

```{r}
f1score_function <- function(df, predicted){
  precision <- precision_function(df, predicted)
  sensitivity <- sensitivity_function(df, predicted)
  f1score <- (2 * precision * sensitivity)/(precision + sensitivity)
  return(f1score)
}

f1score_function(df, predicted)
```


9. Before we move on, let’s consider a question that was asked: What are the bounds on the F1 score? Show
that the F1 score will always be between 0 and 1.

Assume 2*P*S/(P+S)>1 is True
2PS > P+S
2 > 1/S+1/P

Wrong because,
1<1/S and 1<1/P


Assume 2*P*S/(P+S)<0 is True
P, S are positive, f1 score cannot have negative



10. Write a function that generates an ROC curve from a data set with a true classification column (class in our
example) and a probability column (scored.probability in our example). Your function should return a list
that includes the plot of the ROC curve and a vector that contains the calculated area under the curve
(AUC). Note that I recommend using a sequence of thresholds ranging from 0 to 1 at 0.01 intervals.


```{r}
roc_function <- function(df){

  sensitivity_list <- c()
  specificity_list <- c()
  
  for (i in seq(0,1, 0.01)) {

    df$new.scored <- ifelse(df$scored.probability < i, 0, 1)

    sensitivity_list[i*100+1] <- sensitivity_function(df, "new.scored")

    specificity_list[i*100+1] <- specificity_function(df, "new.scored")

  }
  
  roc_df = data.frame(threshold=seq(0,1, 0.01), sensitivity= sensitivity_list, specificity=specificity_list)

  return(roc_df)

}


roc.results <- roc_function(df)

head(roc.results)

ggplot(roc.results, aes(x=1-specificity, y=sensitivity)) + geom_line() + ggtitle("ROC Curve")

```

11. Use your created R functions and the provided classification output data set to produce all of the
classification metrics discussed above.



12. Investigate the caret package. In particular, consider the functions confusionMatrix, sensitivity, and
specificity. Apply the functions to the data set. How do the results compare with your own functions?

```{r}
library(caret)
confusionMatrix(df$scored.class, df$class, positive = "1")

sensitivity_function(df)
specificity_function(df)

```
They match well


13. Investigate the pROC package. Use it to generate an ROC curve for the data set. How do the results

```{r}
library(pROC)

roc_curve <- roc(factor(class) ~ scored.probability, data=df)

plot(roc_curve, col="red", main="ROC curve")

```
