---
title: "HW5_nathanlim"
author: "NathanLim"
date: "July 11, 2016"
output: html_document
---
```{r}
library(dplyr)
library(VIM)
library(ggplot2)
library(MASS)
library(knitr)
library(faraway)
library(gridExtra)
library(caret)
library(pROC)
library(grid)
library(leaps)
library(caret)
library(pROC)
library(grid)
library(reshape2)
library(vcd)
library(glmnet)
```


#Load data

```{r}
raw_wine_data <- read.csv("https://raw.githubusercontent.com/dsmilo/DATA621/master/HW5/Data/wine-training-data.csv")
wine_df <- dplyr::select(raw_wine_data, -INDEX)
```

#Missing values

```{r}
NA_df <- data.frame(NA_counts = colSums(is.na(wine_df)))

kable(NA_df)


df_varHasNA <- dplyr::select(wine_df, -c(TARGET, FixedAcidity, VolatileAcidity, CitricAcid,Density, LabelAppeal, AcidIndex))

aggr(df_varHasNA, prop = TRUE, numbers = TRUE, sortVars = TRUE, cex.lab = 0.4, cex.axis = par("cex"), cex.numbers = par("cex"))
```

# Check variables
```{r}
density_df <- wine_df 

melted <- melt(density_df)

ggplot(melted, aes(value)) + geom_bar(aes(fill = variable, col = variable), alpha = 0.5, show.legend = FALSE) + facet_wrap(~variable, scale="free")+ ggtitle("Density of Skewed Variables\n")
```

We can see TARGET, labelAppeal, AcidIndex, STARS are showing poisson distribution.



#Add variables showing NA 

```{r}
wine_new <- wine_df

wine_new$ResidualSugar_NA <- factor(ifelse(is.na(wine_df$ResidualSugar), 1, 0))
wine_new$Chlorides_NA <- factor(ifelse(is.na(wine_df$Chlorides), 1, 0))
wine_new$FreeSulfurDioxide_NA <- factor(ifelse(is.na(wine_df$FreeSulfurDioxide), 1, 0))
wine_new$TotalSulfurDioxide_NA <- factor(ifelse(is.na(wine_df$TotalSulfurDioxide), 1, 0))
wine_new$pH_NA <- factor(ifelse(is.na(wine_df$pH), 1, 0))
wine_new$Sulphates_NA <- factor(ifelse(is.na(wine_df$Sulphates), 1, 0))
wine_new$Alcohol_NA <- factor(ifelse(is.na(wine_df$Alcohol), 1, 0))
wine_new$STARS_NA <- factor(ifelse(is.na(wine_df$STARS), 1, 0))


p1 <- ggplot(wine_new, aes(x=ResidualSugar_NA, y=TARGET)) + geom_violin() + geom_jitter(alpha=0.2, size=0.2, col="skyblue") + xlab(colnames(wine_new)[16])
p2 <- ggplot(wine_new, aes(x=Chlorides_NA, y=TARGET)) + geom_violin() + geom_jitter(alpha=0.2, size=0.2, col="skyblue") + xlab(colnames(wine_new)[17])
p3 <- ggplot(wine_new, aes(x=FreeSulfurDioxide_NA, y=TARGET)) + geom_violin() + geom_jitter(alpha=0.2, size=0.2, col="skyblue") + xlab(colnames(wine_new)[18])
p4 <- ggplot(wine_new, aes(x=TotalSulfurDioxide_NA, y=TARGET)) + geom_violin() + geom_jitter(alpha=0.2, size=0.2, col="skyblue") + xlab(colnames(wine_new)[19])
p5 <- ggplot(wine_new, aes(x=pH_NA, y=TARGET)) + geom_violin() + geom_jitter(alpha=0.2, size=0.2, col="skyblue") + xlab(colnames(wine_new)[20])
p6 <- ggplot(wine_new, aes(x=Sulphates_NA, y=TARGET)) + geom_violin() + geom_jitter(alpha=0.2, size=0.2, col="skyblue") + xlab(colnames(wine_new)[21])
p7 <- ggplot(wine_new, aes(x=Alcohol_NA, y=TARGET)) + geom_violin() + geom_jitter(alpha=0.2, size=0.2, col="skyblue") + xlab(colnames(wine_new)[22])
p8 <- ggplot(wine_new, aes(x=STARS_NA, y=TARGET)) + geom_violin() + geom_jitter(alpha=0.2, size=0.2, col="skyblue") + xlab(colnames(wine_new)[23])

grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8, ncol=2)

```
When we look at the plots, NAs do not affect the TARGET. However, when STARS is NA, many of TARGET values are zero.


```{r}
df_starsNA=wine_new[which(is.na(wine_new$STARS)==1), ]
table(df_starsNA$TARGET)
```


# Delete the records including missing value except STARS
```{r}
wine_df[which(is.na(wine_df$STARS)), 'STARS'] = 0
table(wine_df$STARS)

wine_cleaned=data.frame(wine_df[complete.cases(wine_df),])
str(wine_cleaned)
```




# Full Model

```{r}
fullmodel_glm <- glm(TARGET ~., data=wine_cleaned)
summary(fullmodel_glm)


fullmodel_poisson = glm(TARGET ~., data=wine_cleaned, family=poisson)
summary(fullmodel_poisson)
```



#BIC
```{r}
regfit.full=regsubsets(TARGET ~., data=wine_cleaned)
reg.summary <- summary(regfit.full)

par(mfrow=c(1,2))
plot(regfit.full, scale = "bic", main = "Predictor Variables vs. BIC")
plot(reg.summary$bic, xlab="Number of Predictors", ylab="BIC", type="l", main="Best subset Selection using BIC")
which.min(reg.summary$bic)
points(7, reg.summary$bic[7], col="red", cex=2, pch=20)
reg.summary$bic
```

When it has 7 variables, the model has the smallest BIC.
However, bic difference between 4 variable model and 7 variable model is not big. 


#Mallow's $C_p$ 

```{r}
par(mfrow = c(1,2))
plot(regfit.full, scale="Cp", main="Predictor Variables vs. Cp")
plot(reg.summary$cp, xlab="Number of Predictors", ylab="Cp", type="l", main="Best subset Selection using Cp" )

which.min(reg.summary$cp) 
points(8, reg.summary$cp[8], col="red", cex=2, pch=20)
```
When it has 8 variables, the model has the smallest Cp.
However, Cp difference between 4 variable model and 8 variable model is not big. 

So we will choose simpler model, 4 variables.


#GLM Model

```{r}
glm_model_1=glm(TARGET ~ VolatileAcidity+LabelAppeal + AcidIndex + STARS, data=wine_cleaned, family=gaussian)

glm_model_2=glm(TARGET ~ VolatileAcidity + Chlorides + FreeSulfurDioxide + TotalSulfurDioxide + Alcohol + LabelAppeal + AcidIndex + STARS, data=wine_cleaned, family=gaussian)

summary(glm_model_1) # AIC: 29609
summary(glm_model_2) # AIC: 29570
```

All of the variables are significant and its AIC is 29609 and 29570



#Poisson Model- variables 4 to 8
```{r}
poisson_model_1=glm(TARGET ~ VolatileAcidity + LabelAppeal + AcidIndex + STARS, data=wine_cleaned, family=poisson)
poisson_model_2=glm(TARGET ~ VolatileAcidity + TotalSulfurDioxide + LabelAppeal + AcidIndex + STARS, data=wine_cleaned, family=poisson)
poisson_model_3=glm(TARGET ~ VolatileAcidity + FreeSulfurDioxide + TotalSulfurDioxide + LabelAppeal + AcidIndex + STARS, data=wine_cleaned, family=poisson)
poisson_model_4=glm(TARGET ~ VolatileAcidity + Chlorides + FreeSulfurDioxide + TotalSulfurDioxide + LabelAppeal + AcidIndex + STARS, data=wine_cleaned, family=poisson)
poisson_model_5=glm(TARGET ~ VolatileAcidity + Chlorides + FreeSulfurDioxide + TotalSulfurDioxide + Alcohol + LabelAppeal + AcidIndex + STARS, data=wine_cleaned, family=poisson)



summary(poisson_model_1) #AIC: 31723
summary(poisson_model_2) #AIC: 31713
summary(poisson_model_3) #AIC: 31706
summary(poisson_model_4) #AIC: 31701
summary(poisson_model_5) #AIC: 31703
```

All of the variables are significant and its AIC is 31723



##10-fold Cross Validation
```{r, echo = FALSE}
k = 10
set.seed(1306)
folds = sample(1:k, nrow(wine_new), replace = TRUE)

cv.errors1 = matrix(NA, k, 10, dimnames = list(NULL, paste(1:10)))
cv.errors2 = matrix(NA, k, 10, dimnames = list(NULL, paste(1:10)))
cv.errors3 = matrix(NA, k, 10, dimnames = list(NULL, paste(1:10)))
cv.errors4 = matrix(NA, k, 10, dimnames = list(NULL, paste(1:10)))
cv.errors5 = matrix(NA, k, 10, dimnames = list(NULL, paste(1:10)))
cv.errors6 = matrix(NA, k, 10, dimnames = list(NULL, paste(1:10)))
cv.errors7 = matrix(NA, k, 10, dimnames = list(NULL, paste(1:10)))
cv.errors8 = matrix(NA, k, 10, dimnames = list(NULL, paste(1:10)))


for (j in 1:k) {
  
  model1=glm(TARGET ~ VolatileAcidity + LabelAppeal + AcidIndex + STARS, data=wine_cleaned, family=poisson)
  
  model2=glm(TARGET ~ VolatileAcidity + TotalSulfurDioxide + LabelAppeal + AcidIndex + STARS, data=wine_cleaned, family=poisson)
  
  model3=glm(TARGET ~ VolatileAcidity + FreeSulfurDioxide + TotalSulfurDioxide + LabelAppeal + AcidIndex + STARS, data=wine_cleaned, family=poisson)
  
  model4=glm(TARGET ~ VolatileAcidity + Chlorides + FreeSulfurDioxide + TotalSulfurDioxide + LabelAppeal + AcidIndex + STARS, data=wine_cleaned, family=poisson)
  
  model5=glm(TARGET ~ VolatileAcidity + Chlorides + FreeSulfurDioxide + TotalSulfurDioxide + Alcohol + LabelAppeal + AcidIndex + STARS, data=wine_cleaned, family=poisson)
  
  model6=glm(TARGET ~ VolatileAcidity+LabelAppeal + AcidIndex + STARS, data=wine_cleaned, family=gaussian)

  model7=glm(TARGET ~ VolatileAcidity + Chlorides + FreeSulfurDioxide + TotalSulfurDioxide + Alcohol + LabelAppeal + AcidIndex + STARS, data=wine_cleaned, family=gaussian)

    for (i in 1:10) {
      f = wine_new[folds == j, ]

      pred1 = predict(model1, f, id = i)
      cv.errors1[j, i] = mean((wine_new$TARGET[folds == j] - pred1) ^ 2, na.rm = TRUE)

      pred2 = predict(model2, f, id = i) %>% exp()
      cv.errors2[j, i] = mean((wine_new$TARGET[folds == j] - pred2) ^ 2, na.rm = TRUE)

      pred3 = predict(model3, f, id = i) %>% exp()
      cv.errors3[j, i] = mean((wine_new$TARGET[folds == j] - pred3) ^ 2, na.rm = TRUE)
      
      pred4 = predict(model4, f, id = i) %>% exp()
      cv.errors4[j, i] = mean((wine_new$TARGET[folds == j] - pred4) ^ 2, na.rm = TRUE)
      
      pred5 = predict(model5, f, id = i) %>% exp()
      cv.errors5[j, i] = mean((wine_new$TARGET[folds == j] - pred5) ^ 2, na.rm = TRUE)
      
      pred6 = predict(model6, f, id = i) %>% exp()
      cv.errors6[j, i] = mean((wine_new$TARGET[folds == j] - pred6) ^ 2, na.rm = TRUE)
      
      pred7 = predict(model7, f, id = i) %>% exp()
      cv.errors7[j, i] = mean((wine_new$TARGET[folds == j] - pred7) ^ 2, na.rm = TRUE)
  }
  
}


mean.cv.errors1 <- apply(cv.errors1, 2, mean)
mean.cv.errors2 <- apply(cv.errors2, 2, mean)
mean.cv.errors3 <- apply(cv.errors3, 2, mean)
mean.cv.errors4 <- apply(cv.errors4, 2, mean)
mean.cv.errors5 <- apply(cv.errors5, 2, mean)
mean.cv.errors6 <- apply(cv.errors6, 2, mean)
mean.cv.errors7 <- apply(cv.errors7, 2, mean)
mean.cv.errors8 <- apply(cv.errors8, 2, mean)

all.cv.error = data.frame(
mean(mean.cv.errors1),
mean(mean.cv.errors2),
mean(mean.cv.errors3),
mean(mean.cv.errors4), 
mean(mean.cv.errors5),
mean(mean.cv.errors6),
mean(mean.cv.errors7),
mean(mean.cv.errors8)
)
names(all.cv.error) = c("Possion Model 1", "Possion Model 2", "Possion Model 3", "Possion Model 4", "Possion Model 5", "Linear Model 1", "Linear Model 2", "Negative binomial Model 1", "Negative binomial Model 2")

all.cv.error = t(all.cv.error)
names(all.cv.error) = c("Model", "Mean CV Error")
all.cv.error

```